Unshuffling Permutations

### 2015-09-20 article.tex
Abstract
Resolution of the ??

\begin{abstract}
    A permutation is said to be a square if it can be obtained by
    shuffling two order-isomorphic patterns. The definition is intended
    to be the natural counterpart to the ordinary shuffle of words and
    languages. In this paper, we tackle the problem of recognizing square
    permutations from both the point of view of algebra and algorithmic.
    On the one hand, we present some algebraic and combinatorial
    properties of the shuffle product of permutations. We follow an
    unusual line consisting in defining the shuffle of permutations by
    means of an unshuffling operator, known as a coproduct. This
    strategy allows to obtain easy proofs for algebraic and combinatorial
    properties of our shuffle product. On the other hand, we prove that
    recognizing square permutations is \NPC. This property is proved by
    establishing a reduction from the pattern involvement problem and by
    using a containment-free perfect matching criterion to decide if a
    permutation is a square.
    %We exhibit a bijection between square
    %$(213,231)$-avoiding permutations and square binary words.
\end{abstract}

@@

  The above claim will be complement in upcoming Claim~\ref{claim: all (nu_1, nu'_1)-edges}.

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
    we have $x' < a'$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}

>>>>

\begin{abstract}
    A permutation is said to be a square if it can be obtained by
    shuffling two order-isomorphic patterns. The definition is intended
    to be the natural counterpart to the ordinary shuffle of words and
    languages. In this paper, we tackle the problem of recognizing square
    permutations from both the point of view of algebra and algorithmic.
    On the one hand, we present some algebraic and combinatorial
    properties of the shuffle product of permutations. We follow an
    unusual line consisting in defining the shuffle of permutations by
    means of an unshuffling operator, known as a coproduct. This
    strategy allows to obtain easy proofs for algebraic and combinatorial
    properties of our shuffle product. We besides exhibit a bijection
    between square $(213,231)$-avoiding permutations and square binary
    words. On the other hand, we prove that recognizing square
    permutations is \NPC. This property is proved by establishing a
    reduction from the pattern involvement problem and by using a
    containment-free perfect matching criterion to decide if a
    permutation is a square.
\end{abstract}

@@

  The above claim will be complement in upcoming Claim~\ref{claim:all (nu_1, nu'_1)-edges}.

@@

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Lemma~\ref{lemma:matching}),
    we have $x' < a'$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}
